### 2020-04-14 thesis/content/chapter-introduction.tex
corrections

Based on the idea of extending \ac{lta} to graphs, we state the following three research questions which will be answered in this thesis:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Formalization of \ac{lta}:}
		The previous work on \ac{lta} provides only a narrow definition of \ac{lta} for unstructured multiset inputs.
		The question what its essential characteristics are, on a more general level, has not yet been considered.
		Therefore we will provide the terminology and definitions to formally capture what constitutes an \ac{lta} method as opposed to a non-\ac{lta} method.
	\item \textbf{An \ac{lta} interpretation of existing \ac{gcr} methods:}
		Using the generalized \ac{lta} formalization, a selection of state-of-the-art \ac{gcr} approaches will be checked for their compatibility with \ac{lta}.
		By doing so, we answer the question whether and how existing \ac{gcr} methods are able to produce predictions which are explainable by local constituent predictions.
	\item \textbf{Definition of a novel \acs{lta}-inspired \ac{gcr} method:}
		As a follow-up question to the previous one, we ask to which extent the existing \ac{gcr} methods are explainable and what their specific shortcomings are.
		Based on those shortcomings we will propose a novel \acs{lta}-inspired approach to overcome them.
\end{enumerate}

@@

\paragraph{\Cref{sec:related}: \nameref{sec:related}}
In order to answer our three research questions, we begin by with on overview of the previous work regarding \ac{lta} as well as \ac{gcr}.
We begin with a brief description of \ac{lta}.
Then the theoretical foundations of the existing \ac{gcr} approaches are introduced;
one particularly important concept there will be the so-called \acl*{wl} coloring.
Based on the theoretical foundations, we conclude the chapter by describing common graph embedding and \acl{gnn} methods.

\paragraph{\Cref{sec:ltag}: \nameref{sec:ltag}}
In this chapter we answer the first two of our research questions.
In the first step a formal definition of \ac{lta} is provided, which answers the question of what constitutes an \ac{lta} method.
This definition is then used to analyze all \ac{gcr} approaches described in \cref{sec:related} for their compatibility with \ac{lta}.

\paragraph{\Cref{sec:ltd}: \nameref{sec:ltd}}
Based on the \ac{lta} perspective on \ac{gcr} provided in \cref{sec:ltag}, we find that the existing approaches are all limited by their underlying solution to the so-called \ac*{ltd} problem.
To overcome this limitation, and thereby answer our third research question, we propose the general idea of edge-filtered graph convolutions.
To realize this general idea in a specific method, we then proceed by proposing a novel type of \ac*{gcnn}, the so-called \textit{2-\acs*{wl}-\acs{gnn}}, which can serve as the foundation for an edge-filtering-based solution to the \acs*{ltd} problem.
Apart from its uses in the context of \ac{lta}/\acs*{ltd}, we show that 2-\acs*{wl}-\acsp{gnn} have additional theoretical advantages over existing approaches;
this makes them interesting even from a non-\acs{lta} perspective.

>>>>

Based on the idea of extending \ac{lta} to graphs, we state the following three research questions which will be answered in this thesis:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Formalization of \ac{lta}:}
		The previous work on \ac{lta} provides only a narrow definition of \ac{lta} for unstructured multiset inputs.
		The question, what its essential characteristics are on a more general level, has not yet been considered.
		Therefore we will provide the terminology and definitions to formally capture what constitutes an \ac{lta} method as opposed to a non-\ac{lta} method.
	\item \textbf{An \ac{lta} interpretation of existing \ac{gcr} methods:}
		Using the generalized \ac{lta} formalization, a selection of state-of-the-art \ac{gcr} approaches will be checked for their compatibility with \ac{lta}.
		By doing so, we answer the question whether and how existing \ac{gcr} methods are able to produce predictions which are explainable by local constituent predictions.
	\item \textbf{Definition of a novel \acs{lta}-inspired \ac{gcr} method:}
		As a follow-up question to the previous one, we ask to which extent the existing \ac{gcr} methods are explainable via the \ac{lta} perspective and what their specific shortcomings are.
		Based on those shortcomings we will propose a novel \acs{lta}-inspired approach to overcome them.
\end{enumerate}

@@

\paragraph{\Cref{sec:related}: \nameref{sec:related}}
In order to answer our three research questions, we begin with an overview of the previous work regarding \ac{lta} as well as \ac{gcr}.
We begin with a brief description of \ac{lta}.
Then the theoretical foundations of the existing \ac{gcr} approaches are introduced;
one particularly important concept there will be the so-called \acl*{wl} coloring.
Based on the theoretical foundations, we conclude the chapter by describing common graph embedding and \acl{gnn} methods.

\paragraph{\Cref{sec:ltag}: \nameref{sec:ltag}}
In this chapter we answer the first two of our research questions.
In the first step a formal definition of \ac{lta} is provided, which answers the question of what constitutes an \ac{lta} method.
This definition is then used to analyze all \ac{gcr} approaches described in \cref{sec:related} for their compatibility with \ac{lta}, which answers the second research question.

\paragraph{\Cref{sec:ltd}: \nameref{sec:ltd}}
Based on the \ac{lta} perspective on \ac{gcr} provided in \cref{sec:ltag}, we find that the existing \ac{gnn} approaches are all limited by their underlying solution to the so-called \ac*{ltd} problem.
To overcome this limitation, and thereby answer our third research question, we propose the general idea of edge-filtered graph convolutions.
To realize this general idea in a specific method, we then proceed by proposing a novel type of \ac*{gcnn}, the so-called \textit{2-\acs*{wl}-\acs{gnn}}, which can serve as the foundation for an edge-filtering based solution to the \acs*{ltd} problem.
Apart from its uses in the context of \ac{lta}/\acs*{ltd}, we show that 2-\acs*{wl}-\acsp{gnn} have additional theoretical advantages over existing approaches;
this makes them interesting even from a non-\acs{lta} perspective.
